---
title: "Week 5 Coding Lab (Penguin Size Comparisons)"
author: "Jonah Seif"
date: "11/5/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(effsize)
library(palmerpenguins)
library(ggbeeswarm)
library(car)
```


# Part 1. T-tests

Mass comparison: male & female chinstrap penguins

```{r}
chinstraps <- penguins %>% 
  filter(species == "Chinstrap")

## Histograms
ggplot(data = chinstraps, aes(x = body_mass_g))+
  geom_histogram(bins = 10) +
  facet_wrap(~sex)

## QQ-plot
ggplot(data = chinstraps, aes(sample = body_mass_g)) +
  geom_qq()+
  facet_wrap(~sex)

## Get descriptive statistics
chinstraps %>% 
  group_by(sex) %>% 
  summarise(mean_mass = mean(body_mass_g),
            median_mass = median(body_mass_g),
            sd_mass = sd(body_mass_g),
            sample_size = n())
```

Get vectors of body_mass_g for female and male chinstrap penguins separately.

```{r}
chinstraps_f <- chinstraps %>% 
  filter(sex == "female") %>% 
  pull(body_mass_g)

chinstraps_m <- chinstraps %>% 
  filter(sex == "male") %>% 
  pull(body_mass_g)
```


### T-tests, a couple different ways

1. Is there a significant difference in body mass between male and female chinstrap penguins?

```{r}
chinstrap_ttest <- t.test(chinstraps_f, chinstraps_m)

chinstrap_ttest

cohen.d(chinstraps_f, chinstraps_m)
```


Are male chinstrap penguins significantly larger (body mass) than female chinstrap penguins?

This is a *one-sided* test - we are only interested in the upper tail. We will still use the `t.test()` function, but update the default `alternative = ` argument from "two-sided" to "greater" or "less."

```{r}
# "Is mean male mass significantly GREATER than female mass?"
t.test(chinstraps_m, chinstraps_f, alternative = "greater")

# Yields the same statistical outcomes as "Is mean female mass significantly LESS than male mass?" 
t.test(chinstraps_f, chinstraps_m, alternative = "less")
```


## One-way ANOVA

### 1. Make a table of counts by species and sex

```{r}
penguins_count <- penguins %>% 
  count(species, sex)

penguins_count
```

We ask: Is there a significant difference in mean flipper length (mm) between the three species for **female** penguins?

- Null hypothesis: There is no significant difference in mean flipper length between the three groups

- Alternative hypothesis: There **is** a significant difference in mean flipper length between the three groups

### 2. First, look at the distributions for female penguin flipper lengths

```{r}
penguins_f <- penguins %>% 
  filter(sex =="female")

ggplot(data = penguins_f, aes(x = flipper_length_mm)) +
  geom_histogram(aes(fill = species),
                 alpha = 0.5,
                 bins = 15,
                 show.legend = FALSE) +
  facet_wrap(~species, scales = "free")
  
ggplot(data = penguins_f, aes(sample = flipper_length_mm)) +
  geom_qq(aes(color = species),
                 show.legend = FALSE) +
  facet_wrap(~species, scales = "free")
```


### Summary statistics 

Starting from penguins_f, create a summary table that groups by species, and returns the following:
- mean of flipper length 
- sd of flipper length 
- sample size
- variance `var`
- standard error = sd / sqrt(n)

```{r}
penguins_f_summary <- penguins_f %>% 
  group_by(species) %>% 
  summarise(mean_flipper_mm = mean(flipper_length_mm),
            sd_flipper_mm = sd(flipper_length_mm),
            sample_size = n(),
            var_flipper_mm = var(flipper_length_mm),
            se_flipper_mm = sd(flipper_length_mm)/sqrt(n()))

# Importantly, variances are pretty close (largest is < 4x greater than the smallest)
# But we can perform a Levene's Test for equal variances anyway: 

leveneTest(flipper_length_mm ~ species, data = penguins_f)
```

### 4. Make a plot of data with mean and SE indicated: 
- Beeswarm plot with mean +- 1sd.

```{r}
ggplot()+
  geom_beeswarm(data = penguins_f, 
                aes(x = species, y = flipper_length_mm),
                color = "gray70",
                size = 3,
                pch = 16,
                alpha = 0.5)+
  scale_x_discrete(labels = c("Adelie", "Chinstrap", "Gentoo")) +
  geom_point(data = penguins_f_summary,
             aes(x = species, y = mean_flipper_mm),
             size = 4)+
  geom_errorbar(data = penguins_f_summary,
                aes(x = species,
                    ymax = mean_flipper_mm + sd_flipper_mm,
                    ymin = mean_flipper_mm - sd_flipper_mm),
                color = "firebrick",
                width = 0.1)+
  theme_light()
```


### One-way ANOVA

```{r}
penguin_aov <- aov(flipper_length_mm ~ species, data = penguins_f)
summary(penguin_aov)

# We retain the alternative hypothesis (mean flipper lengths for the 3 species are NOT the same). But are they all different from each other? Or just two? 

TukeyHSD(penguin_aov)

# Conclusion: means for ALL pairwise combinations are significantly different
```

##### Additional: comparing to Type III ANOVA (with car package):
Note: for unbalanced groups, you may want to do Type III ANOVA instead of Type I (they will yield the same results for balanced design, but for very unbalanced groups you probably want Type III).

```{r}
penguin_lm <- lm(flipper_length_mm ~ species, data = penguins_f)
penguin_aov_t3 <- Anova(penguin_lm, type = "III")
penguin_aov_t3
```



Conclusion: Mean flipper lengths differed significantly between Chinstrap (mean +/- se), Gentoo (mean +/- se), and Adelie (mean +/- se) penguins (one-way ANOVA with post-hoc Tukey's HSD, F(df) = ??, *p* < 0.05). 


